Wealth dynamics in the presence of network structure and primitive cooperation

### 2019-07-12 manuscript/analysis.tex

To see how wealth inequality plays out over time, we plot the Gini coefficient
of the population in Figure~\ref{fig:giniHistory}. (The Gini calculation is
taken with respect to each agent's \textit{effective} wealth, not merely its
personal wealth.) In stage 2, we see inequality decrease as agents join one
another in protos and contribute their ``excess'' wealth to a shared pool. When
the stage 3 starvation period commences, inequality rises as the poorer agents
drop to zero and begin to deplete proto balances.

>>>>

% Will's paragraph goes here, references Figure~\ref{fig:giniHistory}.

### 2019-07-12 manuscript/draftcssa19.tex

The model is realized in a discrete-time Julia simulation program, in which each particle (agent) is represented as a mutable struct that maintains that agent's state over time. Each agent's state includes its current amount of wealth and which proto (if any) it is a member of. Additionally, using the LightGraphs package~\cite{LightGraphs-2017}, each agent is associated with a node of a randomly generated ER graph. Protos are also represented as mutable structs, each of which contains a list of member agent IDs and a current wealth balance.

>>>>

The model is realized in a discrete-time Julia simulation program, in which each agent is represented as a mutable struct (a dedicated area of memory whose contents change over time) that maintains its state. An agent's state includes its current amount of wealth and which proto (if any) it is a member of. Additionally, using the LightGraphs package~\cite{LightGraphs-2017}, each agent is associated with a node of a randomly generated ER graph. Protos are also represented as mutable structs, each of which contains a list of member agent IDs and a current wealth balance.

The simulation proceeds as follows. For configurable parameters
$N > 0$,
$\texttt{init\_max} > 0$,
$\texttt{salary} > 0$,
$m > 0$ (metabolic rate),
$t_p > 0$ (the ``proto threshold''),
$\sigma^2 \ge 0$ and
$0 \le \lambda \le 1$:

\begin{enumerate}
\itemsep.1em
\item Create $N$ agents, each with a random initial wealth (uniformly distributed from 0 to \texttt{init\_max}), and related to one another as per a random ER network with parameter $\lambda$.
\item \textbf{Stages 1 and 2\footnote{We define ``Stage 1'' as the period \textit{before} the first proto is formed, ``Stage 2'' as the period \textit{during} which protos are being formed}:} Repeat until all non-isolate agents\footnote{An \textbf{isolate} agent is one with no graph neighbors.} are members of a proto:
    \begin{enumerate}
    \itemsep.1em
    \item Each agent $A$ whose current wealth $\ge t_p$, and who is not currently a member of a proto, chooses at random one of its graph neighbors (call it $B$) whose wealth \textit{also} exceeds $t_p$. (If there are no such neighbors, proceed to the next agent.) If $B$ is already in a proto, have $A$ join $B$'s proto. If not, have $A$ and $B$ form a new proto.
    \item Each agent gains an amount of wealth equal to $(\texttt{salary} - m + \epsilon)$, where $\epsilon \sim \mathcal{N}(0,\sigma^2)$ (white noise).\footnote{Note that this ``gain'' could be negative, in which case the agent, and possibly its proto, may be subject to death exactly as in step (3).}
    \item Each agent \textit{that is in a proto} donates all its wealth in excess of $t_p$ to that proto's balance. (Agents not in a proto maintain their current wealth.)
    \end{enumerate}
    \item \textbf{Stage 3 (starvation):} Repeat until all agents are dead:
        \begin{enumerate}
        \itemsep.1em
        \item Each agent \textit{loses} an amount of wealth equal to $(m + \epsilon), \epsilon \sim \mathcal{N}(0,\sigma^2)$.
        \item If an agent's wealth would drop below zero as a result of this loss, and if it is not a member of a proto, it dies and is removed from the simulation. If it \textit{is} a member of a proto, it withdraws the necessary amount from its proto's balance to remain at zero wealth. If the proto does not have sufficient funds to cover the loss, both the agent and the proto die and are removed from the simulation.
        \end{enumerate}
\end{enumerate}

Various statistical counters are updated as the program executes so that its behavior can be analyzed post mortem. The main simulation loop can also be invoked from a ``parameter sweep'' program which executes it multiple times over a range of parameter values, in order to determine how the model's behavior changes in response to key parameters.
